---
title: "Census Data Analysis"
output: html_notebook
---

# Overview

Predict whether income exceeds $50K/yr based on census data. Also known as "Adult" dataset.

* 1994 Study
* [Tensorflow](https://www.tensorflow.org/tutorials/wide)
* [Data source](https://archive.ics.uci.edu/ml/datasets/Census+Income)
* [Data description](https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.names)

# Outline

* an Rmd for visualizations/EDA of the raw dataset
* then perhaps an Rmd making predictions
* then perhaps a Shiny app making predictions

# Notes
* it would be cool if we used plotly as that would give us some very nice interactive artifcats
* Important: The training in the demo will occur as an async cloudml job rather than something that happens inline in an Rmd.
* The big drivers are: marital status, education, and sex

```{r setup, message=FALSE}
library(tidyverse)
library(plotly)
```

# Download

```{r, eval=FALSE, message=FALSE}
download.file("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data",
              "data/train_raw.csv")
download.file("https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.test", 
              "data/test_raw.csv")
```

```{r, eval=FALSE, include=FALSE}
# Alternative download location
#download.file("http://mlr.cs.umass.edu/ml/machine-learning-databases/adult/adult.data", "data/train_raw.csv")
#download.file("http://mlr.cs.umass.edu/ml/machine-learning-databases/adult/adult.test", "data/test_raw.csv")
```

# Read Data

```{r, message=FALSE, echo=FALSE, warning=FALSE}
col_names = c(
  "age", "workclass", "fnlwgt", "education", "education_num",
  "marital_status", "occupation", "relationship", "race", "gender",
  "capital_gain", "capital_loss", "hours_per_week", "native_country",
  "income_bracket"
)

train_raw <- read_csv("data/train_raw.csv", col_names = col_names, na = "?")
test_raw  <- read_csv("data/test_raw.csv", col_names = col_names, na = "?", skip = 1)
```

# Create modeling data

```{r}
format.rawdata <- function(data){
  data %>%
    mutate(label = ifelse(income_bracket == ">50K" | income_bracket == ">50K.", 1, 0)) %>%
    mutate(age_buckets = cut(age, c(16, 18, 25, 30, 35, 40, 45, 50, 55, 60, 65, 90))) %>%
    select(label, gender, native_country, education, education_num, occupation, workclass, marital_status, 
           race, age_buckets) %>%
    na.omit
}

create.interactions <- function(data){
  data %>%
    mutate(education_x_occupation = paste(education, occupation, sep = ":")) %>%
    mutate(age_buckets_x_education_x_occupation = paste(age_buckets, education, occupation, sep = ":"))
}

train <- train_raw %>% format.rawdata #%>% create.interactions
test  <- test_raw  %>% format.rawdata #%>% create.interactions
```

# Plot categorical columns

* workclass
* education
* marital_status
* occupation
* relationship
* race
* gender
* native_country

```{r}
plot.main.effects <- function(data, x, y){
  p <- data %>%
    mutate_(group = x, metric = y) %>%
    group_by(group) %>%
    summarize(percent = 100 * mean(metric)) %>%
    ggplot(aes(x = reorder(group, percent), percent)) +
    geom_bar(stat="identity") +
    coord_flip() +
    labs(y = "Percent", x = "") +
    ggtitle(paste("Percent surveyed with incomes over $50k by", x))
  ggplotly(p)
}

plot.main.effects(train, "marital_status", "label")
plot.main.effects(train, "gender", "label")
plot.main.effects(train, "education", "label")
```

# Plot continuous columns

* age
* education_num
* capital_gain
* capital_loss
* hours_per_week

```{r}
plot.continuous <- function(data, x, y, alpha = 0.2, ...){
  data %>%
    select_(groups = x, y = y) %>%
    na.omit %>%
    ggplot(aes(y, fill = groups)) + geom_density(alpha = alpha, ...) +
    labs(x = y, y = "") +
    ggtitle(paste(y, "by", x))
}

# People who earn more also work more, are better educated, and are older
plot.continuous(train_raw, "income_bracket", "age")
plot.continuous(train_raw, "income_bracket", "education_num", adjust = 5)
plot.continuous(train_raw, "income_bracket", "hours_per_week", adjust = 5)

```


# Plot interactions

```{r}
p <- train %>%
  select(education_num, age_buckets, label) %>%
  group_by(age_buckets, education_num) %>%
  summarize(percent = 100 * mean(label)) %>%
  ggplot(aes(education_num, age_buckets, fill = percent)) +
  geom_tile() +
  labs(x = "Education", y = "Age") +
  ggtitle("Percent surveyed with incomes over $50k by age, education")
ggplotly(p)

p <- train %>%
  select(age_buckets, education_num, occupation, label) %>%
  group_by(age_buckets, education_num, occupation) %>%
  summarize(percent = 100 * mean(label)) %>%
  ggplot(aes(education_num, age_buckets, fill = percent)) +
  geom_tile() +
  facet_wrap( ~ occupation) +
  labs(x = "Education", y = "Age") +
  ggtitle("Percent surveyed with incomes over $50k by age, education, and occupation")
ggplotly(p)

```

# Save

```{r}
write_csv(train, "data/train.csv")
write_csv(test, "data/test.csv")
```

